HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2023-02-01 tmlr_submission/hermes_paper_submission/hermes_paper.tex
Minor updates in remaining sections

\section{The HERMES approach: a new hybrid model for time series forecasting}
\label{sec:hybrid}
We introduce a new hybrid approach for time series forecasting  composed of two parts: a collection of per-time-series parametric models, and a global error-corrector neural network train on all time series. Per-time-series parametric models are used to learn local behaviours, to normalize sequences by removing trends and seasonality,  and to compute a first forecast. Then, gathering information of the first predictions and external variables, a recurrent neural network is trained to correct the predictions provided by the first collection of per-time-series models.

@@

\subsection{Per-time-series predictors}
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
\begin{equation}
    \label{eq:predictors}
    \tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
\end{equation}
During the M4 competition, the hybrid model of \citet{smyl2020} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \citet{smyl2020} and the second one uses a TBATS model \citep{alysha2011} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.

In the specific application of fashion trends prediction, two main forms of standard weaknesses were detected and defined for univariate parametric forecasts. The first one is a weakness of definition, when the model is not well defined for its forecasting task. For instance, an additive model provides good predictions on time series with additive seasonality but fails at handling multiplicative seasonalities. The second form, harder to detect and correct, is the weakness of information. For non stationary time series, huge changes of behaviours are not always predictable using the past of the sequence. In some cases, these changes depend on external variables not considered by univariate parametric models. The difficulty is that the exact influence of external variables on the main signal is mostly unknown. This motivates the introduction of a global RNN trained on all time series and able to consider and leverage external signals. First forecast example of our new hybrid model is given in Figure~\ref{fig:introexamples}.

@@

Using well-fitted time-series-specific parametric models, the new hybrid network corrects the first form of weakness and provides appealing performance on the fashion dataset, see Table~\ref{tab:metricresults}. Then, to correct the second form of weakness, in addition to the $N$ target time series, $K \times N$ external sequences indexed from $0$ to $T$ are considered. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ associated with the sequence $\fullts^n$. Let  $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ be all the values of the weak signals associated with the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the sequence. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, a new input for the RNN is defined:   
$$
\fullconcatinput^n_T = \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
$$
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ the final prediction becomes:

>>>>

\section{HERMES: a new hybrid model for time series forecasting}
\label{sec:hybrid}
We introduce a new hybrid approach for time series forecasting  composed of two parts: a collection of per-time-series parametric models, and a global error-corrector neural network train on all time series. Per-time-series parametric models are used in particular to learn local behaviours and to normalize sequences by removing trends and seasonality. Then,  a recurrent neural network driven by the weak signals is trained to correct these per-time-series models.

@@

\subsection{Per-time-series predictors}
%The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. 
For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
\begin{equation}
    \label{eq:predictors}
    \tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
\end{equation}
During the M4 competition, the hybrid model of \citet{smyl2020} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \citet{smyl2020} and the second one uses a TBATS model \citep{alysha2011} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.

%In the specific application of fashion trends prediction, two main forms of standard weaknesses were detected and defined for univariate parametric forecasts. The first one is a weakness of definition, when the model is not well defined for its forecasting task. For instance, an additive model provides good predictions on time series with additive seasonality but fails at handling multiplicative seasonalities. The second form, harder to detect and correct, is the weakness of information. 
For non stationary time series, huge changes of behaviours are not always predictable using the past of the sequence. In some cases, these changes depend on external variables not considered by univariate parametric models. The difficulty is that the exact influence of external variables on the main signal is mostly unknown. This motivates the introduction of a global RNN trained on all time series and able to consider and leverage external signals. First forecast example of our new hybrid model is given in Figure~\ref{fig:introexamples}.

@@

%Using well-fitted time-series-specific parametric models, the new hybrid network corrects the first form of weakness and provides appealing performance on the fashion dataset, see Table~\ref{tab:metricresults}. 
%Then, to correct the second form of weakness, 
In addition to the $N$ target time series, $K \times N$ external sequences indexed from $0$ to $T$ are considered. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ associated with the sequence $\fullts^n$. Let  $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ be all the values of the weak signals associated with the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the sequence. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, a new input for the RNN is defined:   
$$
\fullconcatinput^n_T = \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
$$
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ the final prediction becomes:
